%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Arsclassica Article
% LaTeX Template
% Version 1.1 (1/8/17)
%
% This template has been downloaded from:
% http://www.LaTeXTemplates.com
%
% Original author:
% Lorenzo Pantieri (http://www.lorenzopantieri.net) with extensive modifications by:
% Vel (vel@latextemplates.com)
%
% License:
% CC BY-NC-SA 3.0 (http://creativecommons.org/licenses/by-nc-sa/3.0/)
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%----------------------------------------------------------------------------------------
%	PACKAGES AND OTHER DOCUMENT CONFIGURATIONS
%----------------------------------------------------------------------------------------

\documentclass[
11pt, % Main document font size
a4paper, % Paper type, use 'letterpaper' for US Letter paper
oneside, % One page layout (no page indentation)
%twoside, % Two page layout (page indentation for binding and different headers)
headinclude,footinclude, % Extra spacing for the header and footer
BCOR5mm, % Binding correction
]{scrartcl}

\input{structure.tex} % Include the structure.tex file which specified the document structure and layout

\hyphenation{Fortran hy-phen-ation} % Specify custom hyphenation points in words with dashes where you would like hyphenation to occur, or alternatively, don't put any dashes in a word to stop hyphenation altogether

%----------------------------------------------------------------------------------------
%	TITLE AND AUTHOR(S)
%----------------------------------------------------------------------------------------

\title{\normalfont\spacedallcaps{Designing a Molecular 2-Tag System}} % The article title

%\subtitle{Subtitle} % Uncomment to display a subtitle

\author{\spacedlowsmallcaps{Nicolas Ochsner, Miroslav Phan}} % The article author(s) - author affiliations need to be specified in the AUTHOR AFFILIATIONS block

\date{} % An optional date to appear under the author(s)

%----------------------------------------------------------------------------------------

\begin{document}

%----------------------------------------------------------------------------------------
%	HEADERS
%----------------------------------------------------------------------------------------

\renewcommand{\sectionmark}[1]{\markright{\spacedlowsmallcaps{#1}}} % The header for all pages (oneside) or for even pages (twoside)
%\renewcommand{\subsectionmark}[1]{\markright{\thesubsection~#1}} % Uncomment when using the twoside option - this modifies the header on odd pages
\lehead{\mbox{\llap{\small\thepage\kern1em\color{halfgray} \vline}\color{halfgray}\hspace{0.5em}\rightmark\hfil}} % The header style

\pagestyle{scrheadings} % Enable the headers specified in this block

\tikzset{mytext/.style={rectangle, draw=none, align=center}}

%----------------------------------------------------------------------------------------
%	TABLE OF CONTENTS & LISTS OF FIGURES AND TABLES
%----------------------------------------------------------------------------------------

\maketitle % Print the title/author/date block

\setcounter{tocdepth}{2} % Set the depth of the table of contents to show sections and subsections only

\tableofcontents % Print the table of contents

% \listoffigures % Print the list of figures

% \listoftables % Print the list of tables

%----------------------------------------------------------------------------------------
%	ABSTRACT
%----------------------------------------------------------------------------------------

\section*{Abstract} % This section will not appear in the table of contents due to the star (\section*)

Models of computation are theoretical models, that are used to prove what kind
of computations are possible. One of the most studied models is the universal
Turing machine, which largely defines what computation means nowadays. Another
model, the 2 tag system, can be used to emulate a universal Turing machine. It
works on a string by removing symbols from the front and conditionally appending
symbols at the end. Design of a 2 tag system as a biomolecular process allows us
to compute any problem that can be solved by a universal Turing machine. We
propose an implementation which uses DNA ligation and restriction enzymes to
compute a program that is implemented in a set of engineered sequences.

%----------------------------------------------------------------------------------------
%	AUTHOR AFFILIATIONS
%----------------------------------------------------------------------------------------

%\let\thefootnote\relax\footnotetext{* \textit{Department of Biology, University of Examples, London, United Kingdom}}

%\let\thefootnote\relax\footnotetext{\textsuperscript{1} \textit{Department of Chemistry, University of Examples, London, United Kingdom}}

%----------------------------------------------------------------------------------------

\newpage % Start the article content on the second page, remove this if you have a longer abstract that goes onto the second page

%----------------------------------------------------------------------------------------
%	INTRODUCTION
%----------------------------------------------------------------------------------------

\section{Introduction}

An $m$ tag system is a computational model, that operates on a string of symbols.
At each step it reads the first symbol, removes $m$ symbols from the HEAD of the
string and then appends to its TAIL a sequence of symbols based on the symbol it read to the
string. Mathematically the tag system can be described by the number of symbols
that are removed $m$, the alphabet $\Sigma$ of symbols used and the production
rules, $\Sigma \to \Sigma*$.

\subsection{Example: multiplying two powers of two}
An example implementation of a 2-tag system is multiplying two powers of two, whereby the initial input string consists of \textbf{x} symbols $X$, and \textbf{y} symbols $Y$, where \textbf{x} and \textbf{y} are powers of 2. The string should be converted to $\mathbf{x}\times \mathbf{y}$ symbols $Z$ Threfore to calculate $2^1\times2^2$, then $\mathbf{x}=2,\ \mathbf{y}=4$, the initial string will be $XXYYYY$ and final output will be $ZZZZZZZZ$.

\begin{lstlisting}[
  escapeinside={(*}{*)},
  basicstyle=\ttfamily\small
  ]
  
  2-tag system
  Alphabet: {X, Y, Z}
  Production rules:
     X -> X
     Y -> YZYZ
     Z -> ZZ

     Calculating (*$2\times4$*):
     XXYYYY
       YYYYX
         YYXYZYZ
           XYZYZYZYZ
             XYZYZYZYZ
               ZYZYZYZX
                 ZYZYZXZZ
                    ZYZYZZZZ
                      ZYZZZZZZ
                        ZZZZZZZZ
   
\end{lstlisting}


In the following sections we will first propose a general outline of an
implementation of a 2 tag system and discuss it's features and limitations based
on the processes involved and earlier iterations (namely the iteration we
presented during the last lecture).

\section{Biological Tag System}

\begin{note}
  Unfortunately it turns out that the design of a biological tag system is not
  very easy. (Who would have thought...) However as we still consider it quite
  an interesting idea, we decided to design one for the purpose of discussing
  the solution, its problems and workarounds and the requirements to actually
  make it work.
\end{note}

As an alphabet we use short sequences of DNA which encode a single symbol each.
Our input consists of a sequence of these encodings, where the first symbol
forms a sticky end and the end of the sequence is also exposed as a sticky end.
The general idea is to recognize the first symbol and then conditionally append
the desired production sequence.

This poses the first big design challenge. On one end of our input we have a
specific sticky end, but on the other end we have an unspecific one. Depending
on the one on the front, we need to make sure we only append the correct
sequence. Naturally two basic approaches come to mind.

\begin{enumerate}
  \item Produce or activate the production sequence after detection
  \item Form a plasmid, where the production rule is a sequence connecting both
    ends
\end{enumerate}

\subsection{Plasmid Formation}

Forming a plasmid by connecting the two ends of the input sequence with
production rule sequences seems very elegant at a first glance. The information
is transfered from beginning to end of the sequence by a single fitting
sequence. However, we can not ensure that plasmids properly ligate. It is
possible, even very likely, that two different production sequences ligate to
each end of the input sequence, and thus break the computation.

A program can be encoded as a set of production rules. Each production rule
contains the following four parts.

\begin{enumerate}
  \item Recognition site: A sticky end consisting of four nucleotides encoding
    the symbol to be recognized.
  \item Restriction target: Two restriction enzyme targets. The first one
    deletes two symbols and opens a sticky end for recognition of the next
    read out. The second one prepares a sticky end on the current string to
    append the next tag.
  \item Production sequence: The sequence of symbols or tag, that needs to be
    appended. It only depends on the recognition site.
  \item Terminal site: A unique terminal sequence of length four that can be cut
    into a sticky end and support appending a sequence.
\end{enumerate}

Once the recognition site and the terminal site are both ligated, we add
restriction enzymes. There are two restriction enzyme targeting site. The first
one deletes the first two symbol and creates a new recognition site. The second
one creates a new terminal site. This leaves an input string where we have
conditionally appended a production sequence. After removing the restriction
enzymes we can reproduce the production rules and continue with the next step.

By opening up the sticky sequence after ligation of the recognition symbol, the
problem, where production sequences form chains rather that plasmids can be
solved. This could possibly be done using CRISPR/Cpf1. CRISPR/Cpf1 creates a 5'
overhang of length 4, 21 nt upstream from the PAM. The required PAM is 4 nt
long and would need to be contained between every symbol. Then a properly
constructed guide RNA can detect and cut. Putting a PAM between every symbol
causes sequences to be very long. Thus the cut distance of of CRIPR/Cpf1, 21
nt, is not sufficient to fit the recognition site, restriciton enzyme binding
sites and the production rule inbetween.

Since we were unable to find a set of cutting mechanisms that cut suffitiently
far away, we continue with the design and argue that for slow ligation and
energetically favorable plasmid formation, the error rate may be suffiently
small for short computations. 

%% \subsubsection{Example}

%% To illustrate this design, we give an example encoding of a program that
%% multiplies two powers of two.

%% The three symbols, X, Y and Z can be encoded as

%% \begin{center}
%%   \begin{tikzpicture}
%%     \node[mytext]  (letX)    {X:};
%%     \node[mytext, right of=letX]  (seqX)   {\ttfamily 5-AA-3\\\ttfamily 3-TT-5};
%%     \node[mytext, right of=seqX]  (letY)    {Y:};
%%     \node[mytext, right of=letY]  (seqY)   {\ttfamily 5-GG-3\\\ttfamily 3-CC-5};
%%     \node[mytext, right of=seqY]  (letZ)    {Z:};
%%     \node[mytext, right of=letZ]  (seqZ)   {\ttfamily 5-CC-3\\\ttfamily 3-GG-5};
%%   \end{tikzpicture}
%% \end{center}
%% Thus the initial input of a program that multiplies 2 by 2, XXYY, is

%% \begin{center}
%%   \begin{tikzpicture}
%%     \node[mytext] (seq) {
%%       \ttfamily 5-AATTTAATTTGGTTTGGTTT-3\\
%%       \ttfamily 3-\ \ \ \ ATTAAACCAAACCA\ \ -5
%%     };
%%   \end{tikzpicture}
%% \end{center}
%% The production rules are
%% \begin{center}
%%   \begin{tikzpicture}
%%     \node[mytext] (prodX) {X $\to$ X:};
%%     \node[mytext, right=of prodX] {
%%       \ttfamily 5-AATT\textcolor{cyan}{GCGATG}\textcolor{blue}{GGCGGA}\textcolor{magenta}{(N)}AAAAAAA-3\\
%%       \ttfamily 3-\ \ \ \ \textcolor{cyan}{CGCTAC}\textcolor{blue}{CCGCCT}\textcolor{magenta}{(N)}TTTTT\ \ -5
%%     };
%%     \node[mytext, below=of prodX] (prodY) {Y $\to$ YZ:};
%%     \node[mytext, right=of prodY] {
%%       \ttfamily 5-AACC\textcolor{cyan}{GCGATG}\textcolor{blue}{GGCGGA}\textcolor{magenta}{(N)}AAACCAAAGGAA-3\\
%%       \ttfamily 3-\ \ \ \ \textcolor{cyan}{CGCTAC}\textcolor{blue}{CCGCCT}\textcolor{magenta}{(N)}TTTGGTTTCC\ \ -5
%%     };
%%     \node[mytext, below=of prodY] (prodZ) {Z $\to$ ZZ:};
%%     \node[mytext, right=of prodZ] {
%%       \ttfamily 5-AAGG\textcolor{cyan}{GCGATG}\textcolor{blue}{GGCGGA}\textcolor{magenta}{(N)}AAAGGAAAGGAA-3\\
%%       \ttfamily 3-\ \ \ \ \textcolor{cyan}{CGCTAC}\textcolor{blue}{CCGCCT}\textcolor{magenta}{(N)}TTTCCTTTCC\ \ -5
%%     };
%%   \end{tikzpicture}
%% \end{center}
%% where the letters N correspond to padding (magenta), the cyan sequence is the
%% FokI target site, the blue sequence is the EciI target site, and the S represent
%% the possible terminal sequences.

\subsection{Two Steps}

After reviewing the plasmid formation, we revisited the two step protocol. Using
the idea of detecting ligation of the recognition site using CRISPR, a cascade
can be activated that results in the preparation of the correct production rule.
Instead of using CRISPR/Cpf1, we can use a modified version of Cas9, dCas9, that
allows promotion of transcription. Using dCas9 and a properly constructed sgRNA
allows for conditional transcription of another sgRNA, that can then be used to
activate the corect production rule using a CRISPR/Cpf1 complex to produce a
terminal end.

In this method small nucleotide sequences of length 4 can be used to encode the
symbol. The program can then be defined by two parts:
\begin{enumerate}
  \item Detection sequence: It has recognition site, a restriction enzyme
    target, a buffer sequence followed by the PAM (GG) followed by more buffer
    and a Promoter for an sgRNA that is encoded. Transcription of the sgRNA can
    be activated after ligation of the recognition site. The buffer between the
    restriction enzyme and PAM is constructed such that the sgRNA activating
    transcription overlaps the ligation site. The restriction enzyme target is
    used to prepare the input for the next computational step. It deletes two
    symbols and prepares the next recognition site.
  \item Production sequence: It encodes the symbols to be appended. It also
    provides a target for the sgRNA encoded by the detection sequence, such that
    a sticky end can be created when the sgRNA is present. Further is requires a
    restriction enzyme target, to prepare the input sequence for the next
    computational step after the production sequence has been ligated to the
    input sequence.
\end{enumerate}
Suited type IIS restriction enzymes cut no further than 10 nt from their target site.
Since we need to use them to delete the symbols and we require 2 nt PAMs
inbetween all symbols, we can only use 3 nt for symbol encoding.

\subsubsection{Detection Sequence}

From 3' to 5' the detection sequence starts with the sticky end, that is the
recognition site, directly followed by a buffer nucleotide and a target sequence
for BtgZI (GTAGCG). After that we require the correct amount of spacer
nucleotides before the PAM. The entire sequence, including the recognition site,
can be targeted by sgRNA, that activates transcription of sgRNA.

BtgZI creates a 5' overhang of length 4, 10 nucleotides downstream from its
target site. Once added / activated, it will delete two symbols and create a new
recognition site.

\subsubsection{Production sequence}

From 5' to 3' we have the sgRNA for CRISPR/Cpf1, we then have the terminal site,
which will be opened by Cpf1, followed by the sequence of symbols that needs to
be appended. Each spaced out by CC, which is the PAM required for detection. At
last we have a buffer of one nt and a target sequence for FokI (GTAGG), which
generates a new terminal site. BsaI cuts one nucleotide from its target site and
creates a 5' overhang of size 4.

Note that the terminal site contains two nt that are unique for each symbol. We
thus need a production sequence for every possible terminal sequence. 

\subsubsection{Process outline}

Following there is a step by step outline of the processes involved, describing
the roles of the different sequences and their interaction with each other.
\begin{enumerate}
  \item Detection sequence binds to first symbol in the input sequence and
    ligates to it.
  \item CRISPR/dCas9 is guided to the detection sequence and promotes
    transcription sgRNA specific to the current symbol
  \item CRISPR/Cpf1 is guided to a corresponding production sequence and creates
    a sticky end for the terminal site.
  \item Activated production sequences can now bind to the end of the input
    string
  \item Free detection and production sequences need to be washed out
  \item Restriction enzymes can can be added and target their targeting sites,
    to prepare the input string for the next computational step.
  \item Restriction enzymes need to be removed or knocked out before addition of
    the detection and production sequences
\end{enumerate}

\subsubsection{Example}

%----------------------------------------------------------------------------------------
% RESULTS AND DISCUSSION
%----------------------------------------------------------------------------------------

\section{Discussion}

\subsection{Input length}

The input length can not be smaller that 1, because then we are unable to append
another string. However the input may be arbitrary long.

\subsection{Unique Connections}

Using sgRNA to detect ligation, requires the use of $n$ sgRNAs for $n$ symbols
in the alphabet. Further, since we have $n$ possible terminal sites, we also
need $n$ different versions of the same production rule. For each of these $n$
versions we can use the same activation sgRNA.

If we use the symbols as recognition sites and terminal sites, we can not
entirely avoid formation of plasmids by the input strings. This could be solved
by using a 3' overhang on the recognition site while using a 5' overhang on the
terminal site. However we could not find enzymes that create a 3' overhang far
enough from their target site. This can be solved by reducing the number of
encoding nucleotides to 3 and alternating between having a buffer nucleotide on
one or the other side to avoid plasmid formation. This however means that two
input sequences would not be possible anymore.

\subsection{Remaining Problems}

Ideally we would like to avoid washing out anything. It makes the computations
time and resource expensive. However, then we are unable to ensure activation of
the right components at the right time. It would be thinkable to have some sort
of clock that causes degration and/or inactivation of certain elements, but due
to the many parts involved this is more of a thought than actually realistic
application, because there are enzymes as well as specific sequences involved.
The complexity that emerges from a design like that would certainly be beyond
controlability.

Washing out everything that may interfere with each other creates disadvantages,
but beyond that it is also rather challenging. Specifically, we need to make
sure we do not wash out the input string at any time. While we would be able to
distinguish between used and unused elements, we can not properly distinguish
between the recognition sequences and the input string while solving all other
problems that we encountered.

Further it is unclear if the detection using CRISPR/dCas9 activation is actually
feasible. There might be a lot of leakage caused by the promoter that gets
activated, which will then cause wrong computations to accumulate over time.

The focus of our design is clearly on conditionally appending a sequence
depending on the first symbol. There are several problems that we were unable to
solve completely. While some problems can be solved by complicating the system
and adding more and more control mechanisms, there is the inherent problem of
trying to make a stochastic system behave deterministically. The classical DNA
computing machanisms use this characteristic to explore many solutions at the
same time, while minimizing the number of constraints imposed to achieve a
solution.

The tag system is a way to deterministically formalize the way we think about
computations and mathematically work with it, thus it is not quite suited for
DNA computation. If one wanted to find existing computational models that are
more suited for the description of biological computer, one might find those in
the very advanced area of theoretical computer science, where probabilistic
models are used to better understand the boundaries of approximation algorithms
for NP-hard problems etc.

%----------------------------------------------------------------------------------------
%	BIBLIOGRAPHY
%----------------------------------------------------------------------------------------

\renewcommand{\refname}{\spacedlowsmallcaps{References}} % For modifying the bibliography heading

\bibliographystyle{unsrt}

\bibliography{sample.bib} % The file containing the bibliography

%----------------------------------------------------------------------------------------

\end{document}
